Forecasting the 2024 Election: Trump Dominates the South, Harris Leads in the Midwest

### 2024-10-22 paper/paper.qmd
committing text changes to paper.qmd for model testing

Out-of-sample testing results [Appendix -@sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S. presidential election. Our accuracy test generated a 75% score for the model, indicating our model's high reliability for making predictions about state-level election outcomes. Since presidential elections hinge on state-level wins, a reliable state-by-state prediction would imply a reliable national outcome prediction.

The RMSE value generated for our model is 0.5, which indicates that the model-predicted probabilities are close to the true outcomes. The confusion matrix breaks down the model’s predictions into four categories: 1. correctly predicted Trump wins, 2. correctly predicted Harris wins, 3. Incorrectly predicted Trump wins, 4. Incorrectly predicted Harris wins. The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model. Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix -@sec-model-diagnostics].

@@

## Diagnostics {#-sec-model-diagnostics}

@@

# Confusion Matrix for additional diagnostics
conf_matrix <- table(Predicted = test_pred_binary, Actual = test_data$winner)

@@

[@barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election. The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted. The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states. This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.

```{r}
#| echo: false
#| eval: true
#| label: barplot-diagnostics
#| fig-cap: "Model: Diagnostics: Accuracy and RMSE"
#| warning: false
#| message: false

@@

Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election. To represent our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via [@US-election-survey].

>>>>

Out-of-sample testing results [Appendix @sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S. presidential election. Our accuracy test generated a 75% score for the model, indicating our model's high reliability for making predictions about state-level election outcomes. Since presidential elections hinge on state-level wins, a reliable state-by-state prediction would imply a reliable national outcome prediction.

The RMSE value generated for our model is 0.5, which indicates that the model-predicted probabilities are close to the true outcomes. The confusion matrix breaks down the model’s predictions into four categories: 1. correctly predicted Trump wins, 2. correctly predicted Harris wins, 3. Incorrectly predicted Trump wins, 4. Incorrectly predicted Harris wins. The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model. Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix @sec-model-diagnostics].

@@

## Diagnostics {#sec-model-diagnostics}

@@

# Confusion Matrix for additional diagnostics
conf_matrix <- table(Predicted = test_pred_binary, Actual = test_data$winner)
rownames(conf_matrix) <- c("Predicted 0 (Harris)", "Predicted 1 (Trump)")
colnames(conf_matrix) <- c("Actual 0 (Harris)", "Actual 1 (Trump)")

@@

[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election. The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted. The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states. This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.

```{r}
#| echo: false
#| eval: true
#| label: fig-barplot-diagnostics
#| fig-cap: "Model: Diagnostics: Accuracy and RMSE"
#| warning: false
#| message: false

@@

Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election. To represent our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the following url: https://docs.google.com/forms/d/1S6aN5Q82orjKYJI4i7001sPyn5jhRkgNIkI9teiQq_o/edit [@US-election-survey].
